---
title: "Conditional Probability"
author: "Hans Lehndorff"
contributor: "DATA 505: Applied Machine Learning -- Professor Hendrik Orem"
date-modified: "today"
code-fold: false
---

```{r setup, include=FALSE}
#| echo: false
#| warnings: false
#| messages: false

library(tidyverse)
library(ggplot2)

```

There are three cards. One with two red sides, one with two black sides, one with a black side and a red site.

A random card is placed in front of you -- it has a red face. What are the odds that the other side of the card is black?

Intuitively, I would expect the answer to be 50% -- there are only two possible options red or black. Given that we are not looking at the black/black card the odds that the other side is black (this is the red/black card) and the odds that the other side is red (this is the red/red card) should be equally likely.

This is a simulation of this situation as a means of testing Bayes Theorem for conditional probabilities.

First, there are six possibilities for how the cards could be oriented -- each of the three card with one of two sides showing.

```{r options}
#| echo: true
#| messages: false
#| warnings: false


cards<-data.frame(
  card=1:6,
  face=c("R","R","R","B","B","B"),
  back=c("R","R","B","R","B","B"),
  card_num=c(1,1,2,2,3,3),
  orientation=rep(c("A","B"),3)
  )

cards

```

These cards can then be randomly sampled 1,000,000 times.

```{r samples}
#| echo: true
#| messages: false
#| warnings: false

sampled<-data.frame(card=sample(1:6,1000000,replace=T)) %>% 
  left_join(cards)
```

To calculate the probability that the other side of the card is black given that we are looking at a red side, we first need to filter our random samples down to the cases where a red card is face up. Then count up the frequency with which black and red each appear as the back side of the card.

```{r result, message=FALSE}

sampled %>% 
  filter(face=="R") %>% 
  group_by(back) %>% 
  summarise(n=n()) %>% 
  mutate(pct=scales::percent(n/sum(n)))

```

Rather than being the expected 50% probability, the odds of having the other side be black is only 1/3.
