---
title: "Impact on Electricity Usage Caused by the COVID-19 Pandemic"
author: "Hans Lehndorff"
contributor: "Cleighton Roberts"
date-modified: "today"
---

```{r, include=FALSE}
library(tidyverse)
library(gt)
library(lubridate)

load('data/NEEA HEMS/final_dataset.rdata')
load('data/NEEA HEMS/analysis_data.rdata')
points<-read_csv("data/NEEA HEMS/points.csv")
sites<-read_csv("data/NEEA HEMS/sites.csv")
```

# Study Background

[NEEA HEMS](https://neea.org/data/nw-end-use-load-research-project)

# Data Overview

Data represent a diverse geographic area.

```{r map}
sites_with_mains<-subset(points,circuit_label_type_desc=="Mains")$ee_site_id
these_eus<-c(
  "Mains","Hot Tub","Electric Resistance Storage Water Heaters",
  "Central AC", "Clothes Washer", "Clothes Dryer","Refrigerator/Freezer",
  "Other","Stove/Oven/Range","Electric Vehicle Charger","Dishwasher",
  "Ductless Heatpump","Electric Furnace","Ducted Heatpump"
  )

#Step 1 drop incomplete circuits
circuit_agg<-analysis_data %>% 
  filter(ee_site_id%in%sites_with_mains) %>% 
  group_by(ee_site_id,regname) %>% 
  summarise(n=n())

retain_circuits<-circuit_agg %>% 
  filter(n>12755)

analysis_data<-analysis_data %>% 
  inner_join(retain_circuits %>% select(-n))

#Step 3 drop errant days
day_agg<-analysis_data %>% 
  group_by(date,hour) %>% 
  summarise(n=n())

retain_hours<-day_agg %>% 
  filter(n>625) %>% 
  filter(!date%in%as.Date(c("2019-09-25","2019-09-26")))

analysis_data<-analysis_data %>% 
  inner_join(retain_hours %>% select(-n))

weather_stations <- read_csv("data/NEEA HEMS/weather_stations.csv") %>% 
  mutate(station_id=paste(USAFID,WBAN,sep="-"))

state_maps<-map_data("state",c("oregon","washington","idaho","montana"))

sites_ws<-sites %>% 
  filter(ee_site_id%in%analysis_data$ee_site_id) %>% 
  left_join(
    weather_stations %>% 
      group_by(station_id) %>% 
      filter(1:n()==1)
  ) 

data_for_map<-analysis_data %>% 
  left_join(sites_ws %>% select(ee_site_id,station_id,LAT,LONG)) %>% 
  group_by(station_id,LAT,LONG) %>% 
  summarise(sites=n_distinct(ee_site_id), data=n())

city_labs<-maps::us.cities %>% 
  filter(country.etc%in%c("OR","ID","WA","MT")) %>% 
  filter(name%in%c("Seattle WA","Bend OR","Portland OR","Boise ID","Spokane WA","Missoula MT")) %>% 
  mutate(lat=ifelse(name%in%c("Bend OR"),lat-.5,lat+.5),
         long=ifelse(name%in%c("Seattle WA"),long-2,long))

ggplot()+
  geom_polygon(data=state_maps,aes(x=long,y=lat,group=group),fill='white',color='black')+
  geom_point(data=data_for_map,aes(x=LONG,y=LAT,size=data,fill=sites),color="black",shape=21,alpha=.5)+
  geom_label(data=city_labs,aes(x=long,y=lat,label=name))+
  coord_map(xlim = c(-127,-110))+
  theme_void()+
  scale_size_continuous(labels = scales::comma)+
  scale_fill_viridis_b()+
  theme(
    legend.position = 'bottom',
    legend.box= 'horizontal')+
  labs(fill="Number of Sites",size="Total Observations",
       title="Geographic Distrobution of NEEA HEMS Data",
       subtitle = "Behavioral insights from sample of homes across the Pacific Northwest")+
   guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5),
         size = guide_legend(title.position="top", title.hjust = 0.5))


```

A variety of end uses are represented in the data.

```{r eus}
eu_summary<-analysis_data %>% 
  left_join(points %>% select(ee_site_id,regname,circuit_label_type_desc)) %>% 
  group_by(circuit_label_type_desc) %>% 
  summarise(
    n=n()
  ) %>% 
  filter(!circuit_label_type_desc%in%c("Other","Mains","Gas Furnace (Component)")) %>% 
  mutate(
    category=case_when(
      circuit_label_type_desc%in%c("Central AC","Ducted Heatpump","Ductless Heatpump","Electric Baseboard Heaters","Other Zonal Heat","Electric Furnace") ~ "HVAC",
      grepl("water heater",circuit_label_type_desc,ignore.case = T) ~ "Water\nHeating",
      circuit_label_type_desc%in%c("Stove/Oven/Range","Microwave","Garbage Disposal","Dishwasher","Refrigerator/Freezer") ~ "Kitchen",
      TRUE~"Other"
    )
  )

ggplot(eu_summary)+
  geom_col(aes(x=n,y=forcats::fct_reorder(abbreviate(circuit_label_type_desc,30),n),fill=category))+
  scale_x_continuous(labels = scales::comma,expand=c(0,0,.05,0))+
  scale_fill_brewer(type="qual",palette = 2)+
  theme_classic()+
  facet_grid(category~.,scale="free_y",space="free")+
  labs(x="Number of Observations",y="Equipment Type",fill="Equipment Category",
       title="Appliance Distrobution of NEEA HEMS Data",
       subtitle = "Data available for range of household appliances")+
  theme(
    legend.position = 'bottom',
    strip.text.y = element_text(angle = 0),
    title = element_text(size=9))+
  guides(fill='none')
```

# Summary Statistics

For some end uses, energy usage was considerably different when comparing March to September of 2020 to March to September 2019.

```{r stoves}

plot_data<-final_dataset %>% 
  filter(circuit_label_type_desc=="Stove/Oven/Range") %>% 
  filter(doy<350) %>% 
  group_by(date,date_2020) %>% 
  summarise(kwh=sum(kwh)) %>% 
  arrange(date) %>% 
  group_by(year(date)) %>% 
  mutate(sma=zoo::rollmean(kwh,30,na.pad = T,align = 'right'))

ggplot(plot_data)+
  geom_smooth(aes(x=date_2020,y=kwh,color=factor(year(date))),linewidth=1.5,method = 'loess',span=.17,se=F)+
  geom_hline(yintercept = 0,color=NA)+
  geom_vline(xintercept = as.Date("2020-03-08"))+
  labs(x="Day of the Year",y="Smoothed Daily Energy Usage (kWh)",color="Time Period",
       title="Comparison of Stove Energy Usage by Day of Year", subtitle = "Stove usage increased dramatically when the Pandemic started")+
  scale_color_brewer(type = "qual",palette =  6)+
  scale_x_date(date_breaks = "1 month",date_labels = "%b",limits = c(as.Date("2020-01-01"),as.Date("2020-09-01")))+
  theme_minimal()+
  theme(
    legend.position = 'bottom',
    panel.grid.minor.x = element_blank()
    )

```

# Summary Table

A summary table of each end use including 2019 and 2020 average daily usage (between March 15 and September 1 of each year), percentage change and the distribution of daily usage. Results have not been weather normalized.

```{r gt_table}

initial_table<-final_dataset %>% 
  filter(doy<350&doy>yday("2020-03-15")) %>% 
  group_by(circuit_label_type_desc,date) %>% 
  summarise(kwh=sum(kwh)) %>% 
  group_by(circuit_label_type_desc,Year=year(date)) %>% 
  summarise(
    kWh=round(mean(kwh),3)
    ) %>% 
  pivot_wider(names_from = Year,values_from = kWh) %>% 
  mutate(`Percent Change`=round((`2020`-`2019`)/`2019`,4)) %>% 
  arrange(-abs((`2020`-`2019`)/`2019`)) %>% 
  mutate(Histogram=NA) %>% 
  ungroup()

for(eu in initial_table$circuit_label_type_desc){
  # print(eu)
  plot<-final_dataset %>% 
    filter(doy<350&doy>yday("2020-03-15")) %>% 
    filter(circuit_label_type_desc==eu) %>% 
    group_by(circuit_label_type_desc,date,year=factor(year(date))) %>% 
    summarise(kwh=sum(kwh)) %>% 
    ggplot()+
    geom_density(aes(x=kwh,color=year,fill=year),alpha=.4)+
    scale_color_brewer(type = "qual",palette =  "Set1")+
    scale_fill_brewer(type = "qual",palette =  "Set1")+
    scale_y_continuous(breaks = NULL)+
    theme_minimal()+
    theme(legend.position = 'bottom')+
    labs(x="Daily kWh",y=NULL,fill="Year",color="Year")
  
  initial_table$Histogram[initial_table$circuit_label_type_desc==eu]<-list(plot)
}


output_table <- initial_table %>%
  rename(
    "2019 Average Daily Usage" = `2019`,
    "2020 Average Daily Usage" = `2020`,
    "End Use"=circuit_label_type_desc) %>% 
gt() %>%
text_transform(
 locations = cells_body(Histogram),
 fn = function(Histogram) {
   map(initial_table$Histogram, ggplot_image, height = px(150))
 }
) %>% 
  fmt_percent(columns = `Percent Change`) %>% 
  gtExtras::gt_color_rows(
    columns = `Percent Change`,
    palette = c("#E41A1C","white", "#377EB8"),
    direction = 1,
    domain = c(-1,1),
    pal_type = "continuous"
    )

# brewer.pal(2,"Set1")
output_table


```
