---
title: "What is HCL?"
editor: visual
author:
- name: Hans Lehndorff
date-modified: "today"
---

HCL is a colorspace which is better than RGB for creating perceptually uniform colors.

-   Hue: wavelength (color)

-   Chroma: purity, dullness

-   Luminance: from black to white

The polarLUV function from the colorspace package, takes hue, chroma and luminance inputs and returns a color hexcode (via the hex function). Based on testing, polarLUV:

-   takes hue as a value between -134 and 107

-   luminance of 0 = black

-   negative chroma exist, but I find their interpretation less intuitive

-   there are many impossible combinations of HCL.

This process goes through a range of hues and generates a sample combinations of chroma and luminance.

```{r some_data}
#| code-fold: false
#| echo: true
#| message: false
#| warning: false

library(tidyverse)
library(colorspace)

# -134:107

output<-NULL
for(h in seq(-134,107,16)){
  # print(h)
  z=expand.grid(
  l=seq(1,150,.1),#
  c=seq(-200,200,.1) #
  ) %>% 
    mutate(
      h=h,
      color=hex(polarLUV(H=h,C=c,L=l))
      ) %>% 
    filter(!is.na(color)) %>% #
    slice_sample(n=10000,replace = T) %>% 
    distinct()
  
  output<-bind_rows(output,z)
  
}

```

A plot showing the impact on color of chroma and luminance on a selection of hues.

```{r hcl_plot,fig.height=6, fig.width=10}
ggplot(output)+
  geom_point(aes(x=l,y=c),color=output$color)+
  geom_hline(yintercept = 0, linetype=2)+
  theme_minimal()+
  labs(x="Luminance",y="Chroma",title="Impact of Luminance and Chroma variation for a sample of Hues")+
  facet_wrap(h~.)
```
